---
title: "FindingPath-Windows"
output: html_document
---

README: This file is NOT a necessary component of the Guide Finder program. It is intended to assist users in finding the full PATH for the makeblastdb and blastn files. The PATH can be identified in a number of other ways. This script searches all directories for the makeblastdb 
and blastn files, which should be kept within the GuideFinder-Mac folder. After the print() command, copy the full PATH from the Console and paste into the proper location in the Preprocessing or Guide Finder script. 

##1. Set the Working Directory 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/spotom/Desktop')
```


##2. Find the full PATH for makeblastdb file 
This chunk of code with find the full PATH (e.g.) for the makeblastdb file located within GuideFinder-Windows folder
```{r}

MakeBlastDB<-system('where /r C:\\ makeblastdb.exe', intern = TRUE)
grep("GuideFinder-Windows",MakeBlastDB, value = TRUE)

```

##3. **Copy the full path for the makeblastdb file from the console**



##4. Find the full PATH for blastn file 
```{r}
BlastnLocation<-system('where /r C:\\ blastn.exe', intern = TRUE)
grep("GuideFinder-Windows",BlastnLocation, value = TRUE)

```

##5. **Copy the full path for the blastn file from the console**
