---
title: "FindingPath_Mac"
output: html_document
---

README: This file is NOT a necessary component of the Guide Finder program. It is intended to assist users in finding the full PATH for the makeblastdb and blastn files, located within the downloaded GuideFinder-Mac folder. 
The PATH can be identified in a number of other ways. This script searches all directories for the makeblastdb and blastn files, which should be kept within the GuideFinder-Mac folder. 
After the print() command, copy the full PATH from the Console and paste into the proper location in the Preprocessing or Guide Finder script. 

##1. Set the Working Directory 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/spotom/Desktop')
```


##2. Find the full PATH for makeblastdb file 
This chunk of code with find the full PATH (e.g./Users/spotom/Documents/Temp/Guide-Finder-master/GuideFinder-Mac/makeblastdb) for the makeblastdb file
```{r}
system('find / -name makeblastdb > MakedbLocation.txt ')
system('grep "GuideFinder-Mac" MakedbLocation.txt > MakePathDB.txt')
FullPath_MakeDB<-fread("MakePathDB.txt", header = F)
names(FullPath_MakeDB)<-c("FullPath_MakeDB")
print(FullPath_MakeDB$FullPath_MakeDB) 

```

##3. **Copy the full path for the makeblastdb file from the console**


##4. Find the full PATH for blastn file 
This chunk of code with find the full PATH (e.g./Users/spotom/Documents/Temp/Guide-Finder-master/GuideFinder-Mac/blastn) for the blastn file
```{r}
system('find / -name blastn > BlastLocation.txt ')
system('grep "GuideFinder-Mac" BlastLocation.txt > BlastPath.txt')
BlastPath<-fread("BlastPath.txt", header = F)
names(BlastPath)<-c("BlastPath")
print(BlastPath$BlastPath)
```

##5. **Copy the full path for the blastn file from the console**
