---
title: "PreProcessing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/spotom/Desktop')
setwd('/Users/spotom/Desktop') #set working directory

```


__Install Packages__
These packages need only to be installed once 
```{r, eval=FALSE, echo=FALSE}

```


__Loading Packages__
Packages need to be loaded each time the program is run 
```{r, eval=FALSE, echo=FALSE}
library(BiocInstaller)
library(seqinr)
library(Biostrings)
library(ape)
library(refGenome)
library(genbankr)
library(rentrez)
library(stringr)
library(reshape)
library(plyr)
library(annotate)
library(data.table)
library(RFLPtools)
library(dplyr)

```


__Functions Created__
```{r, eval=FALSE, echo=FALSE}

String2<-function(x) {
  Step<-strsplit(x, split = ". ")[[1]][1]
  strsplit(Step, split = "|", fixed = TRUE)[[1]][2]
}

StrandType<-function(x) { 
  if (x < 0) { 
    print("-")}
  else {
    print("+")}
}


Subtract50<-function(x) {
  x - 50
}

Add50<-function(x) { 
  x + 50 
}

Split1<-function(x) {
  strsplit(x, split = "|", fixed = TRUE)[[1]][2]}

Ns<-"NNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNN" #to add between contigs 

```

__Making a multi-sequence fasta file into a single-sequence fasta file__
This must be the first step in the program, since the remainder of the program requires a single-sequence fasta file as input. 
Making a single-sequence fasta file from a multi-sequence fasta file (with a string of 60Ns between contigs). 
This pre-procesing script and the guide finder program require a single sequence fasta file. Your genome sequence may be in several contigs rather than one sequence 
so this chunk of code concatenates the contigis and adds a series of Ns between the contigs. Substitute the file name for the name of your file. 
Example: Example: fastafile<-read.fasta(file= "Aureus_SA14_Draft_fasta.fasta", as.string = TRUE); replace Aureus_SA14_Draft_fasta.fasta with the name of your file. May be .fasta or .txt file. 
If you already have a single-sequence file, skip this chuck and simply upload the fasta file using the next chunk of code. 

```{r, eval=FALSE, echo=FALSE}
fastafile<-read.fasta(file="staphaureusexamplefasta_multiseq.txt", as.string = TRUE) #read in fasta file

seqnames<-names(fastafile) #subset seq names/IDs
seq<-paste(fastafile) #subset sequences
seqdf<-data.frame(seqnames, seq) #paste together in a data frame 
#loop to add Ns to begining of each contig
NewFasta<-NULL 
for (i in 1:nrow(seqdf)) { 
  seqrow<-seqdf[i,]
  seqrow$newseq<-paste(Ns,seqrow$seq, sep = "", collapse ="")
  newseqs<-cbind( seqrow$seqnames, seqrow$newseq)
  NewFasta<-rbind(NewFasta, newseqs)
}
NewFasta1<-data.frame(NewFasta, stringsAsFactors = FALSE) #make data frame
names(NewFasta1)<-c("ID", "sequences") #rename

SingleSeq<-paste(NewFasta1[1:nrow(NewFasta1),2], collapse = "") #paste together sequences 
SingleSeqFasta<-DNAString(SingleSeq) #make into DNA string object 
write.fasta(SingleSeqFasta,"SingleSequenceFastaFile",file="concatenatedfastafile.txt") #write out to fasta file 
```


__Add file names here__ 
In this chunk of code, add the names of the files you'll be pre-processing.  
For this preprocessing script, you'll need:  
* Coding sequence file: this file contains the nucleotide sequence for each gene in the entire genome in the 5' --> 3' direction 
* Fasta file: this file contains the entire nucleotide sequence of the genome. For draft genomes, this file may be broken up into different  
contigs within the same file. However, the guide finder program requires a single-sequence fasta file as input.  
For this reason, the previous chuck of code has concatenated these chuncks of sequence into one contiguous sequence. __Here, input this new, concatenated sequence__
Add the names of your files here:  
Example:  
InputGenome<-readDNAStringSet( __"concatenatedfastafile.txt"__, use.names = TRUE); replace "concatenatedfastafile.txt"" with the name of your single-sequence file, if you did not create your file with the above chunk of code  
(i.e. you already had a single-sequence file).  
MakeDatabase<-' __/Users/spotom/Desktop/ncbi-blast-2.6.0+/bin/makeblastdb__ -in __concatenatedfastafile.txt__ -dbtype nucl -out GenomeDatabaseEpi'; replace the file path with the file path where you makeblastdb is kept; replace "concatenatedfastafile.txt" if you have a different file than created above.  
CDS<- readDNAStringSet(" __Aureus_CDS.fasta__", use.names = TRUE); replace "Aureus_CDS.fasta" with the name of your coding sequence file.   
RunBlast<-' __/Users/spotom/Desktop/ncbi-blast-2.6.0+/bin/blastn__ -task megablast -query __Aureus_CDS.fasta__ -db GenomeDatabaseEpi -outfmt 6 -num_threads 4 -dust no -soft_masking false -max_target_seqs 1 -perc_identity 100 -out GenesMatch.blast'; replace the file path for the path of blastn. replace "Aureus_CDS.fasta" with the name of your coding sequence file.   


```{r, eval=FALSE, echo=FALSE}
#Input fasta file and change file path for MakeDatabase
InputGenome<-readDNAStringSet("concatenatedfastafile.txt", use.names = TRUE) 
MakeDatabase<-'./GuideFinderMac/makeblastdb  -in concatenatedfastafile.txt -dbtype nucl -out GenomeDatabaseEpi'

#Input CDS file
CDS<- readDNAStringSet("staphaureusexample_codingseq.fasta", use.names = TRUE)
RunBlast<-'./GuideFinderMac/blastn -task megablast -query staphaureusexample_codingseq.fasta -db GenomeDatabaseEpi -outfmt 6 -num_threads 4 -dust no -soft_masking false -max_target_seqs 1 -perc_identity 100 -out GenesMatch.blast'

```


__Making into data frame__
This chunk just creates a data frame from the CDS file, containing sequence information and gene names/IDs  
```{r, eval=FALSE, echo=FALSE}
seqnames<-names(CDS) #getting seq names 
seqnames<-data.frame(seqnames, stringsAsFactors = FALSE) #putting sequence names into DF
CDSDF<-data.frame(CDS, stringsAsFactors = FALSE) #making string set into a data frame
CDSDF<-cbind(CDSDF, seqnames)
names(CDSDF)<-c("CDS", "Name")
row.names(CDSDF)<-c() #removing row names 

```

__Subsetting gene identifiers__  
This is used to subset gene names/IDS from the long name in the CDS file  
__WARNING:__ As geonome annotation notation varies significantly, this section is very unstable. You may need to adjust the code above (in function section) to accurately subset.  
The functions used to subset the names are listed above to subset gene names/identifiers/product information from the coding sequence file.  
The column names that this script assigns ("Gene Product", "Name") are used mostly to be consistent with the main guide finder script.  
For this reason, The information pulled from the file may not appear to be categorized correctly (e.g. "Gene Product" may not be a protein product name but rather a unique alphanumeric identifier). This is OK.
```{r}
CDSDF$UniqueID<-lapply(CDSDF$Name, String2)
#CDSDF$Product<-lapply(CDSDF$Name, GeneProduct)
CDSDF$Name<-c() #removing name column 
CDSDF$ID<-1:nrow(CDSDF) #making IDs for each CDS
CDSDF<-as.data.frame(lapply(CDSDF, unlist))
newdf1<-cbind(paste(CDSDF$UniqueID),CDSDF)
newdf1$UniqueID<-NULL
CDSDF<-newdf1
names(CDSDF)<-c("UniqueID", "CDS", "ID")

```

__Run BLAST__
This chunk runs BLAST to align the coding sequences against the fasta file -- output that we want is the coordinates for each coding sequence.
```{r, echo=FALSE, eval=FALSE}
system(MakeDatabase) #Make blast databse (from nt sequence file)
system(RunBlast) #blast coding sequence against genome to get coordinates (later used to also pick up puative promoter region)
system('sort -u -f -k1,1 GenesMatch.blast > singleBLASThit.txt') #select only one top hit from hits
```

__Subset BLAST output__
This section reads in the blast output and subsets the information to just contain the gene name/ID and the gene start and end coordinates
```{r}
TopHits<-fread("singleBLASThit.txt") #Reading in blast output
TopHits$UniqueID<-TopHits$V1 #add names
TopHits<-as.data.frame(TopHits)
Subset<-c("V9", "V10", "UniqueID") #subset coordinates and names 
SubsetTopHits<-TopHits[Subset]
names(SubsetTopHits)<-c("start", "end", "UniqueID") #rename
```

__Subsetting BLAST output Names__
This section attempts to split the geneproduct/id into a name that matches the geneproduct/id subset earlier; we're going to merge on this name/ID so they need to exactly match
__WARNING__: This section may also be unstable, given varaible genome annotation. May need to adjust Split1 function to be able to accurately subset the GeneProduct names in the dataframe  
SubsetTopHits so that they match exactly the "Gene Product" names in the CDSDF data frame. 
```{r, echo=FALSE, eval=FALSE}
SubsetTopHits$UniqueID<-lapply(SubsetTopHits$UniqueID, Split1) #more splitting
newdf<-cbind(paste(SubsetTopHits$UniqueID), SubsetTopHits)
newdf$UniqueID<-NULL
SubsetTopHits<-newdf
names(SubsetTopHits)<-c("UniqueID","start","end")

```


__Merging data frames__
merging CDSDF with blast hit output to correlate the coordinates identified in the BLAST output with the values in the CDSDF data frame. 
```{r, echo=FALSE, eval=FALSE}
GeneProductsDF<-merge(CDSDF, SubsetTopHits, by = "UniqueID")
GeneProductsDF<-GeneProductsDF[c("UniqueID", "start", "end")]
```

__Adjusting Coordinates__
The first part of this chunk makes the coordinates numeric and identifies whether the gene is coded on the plus or minus strand.  
The loop makes the coordinates into the format that we want (the format that the main guide finder script requires; for genes coded on the minus strand, we want to switch the start and end coordinates)
```{r, echo=FALSE, eval=FALSE}
GeneProductsDF$start<-as.numeric(as.character(GeneProductsDF$start)) #making start coordinate numeric
GeneProductsDF$end<-as.numeric(as.character(GeneProductsDF$end)) #making end coordinate numeric
GeneProductsDF$CoordDiff<-(GeneProductsDF$end - GeneProductsDF$start) #finding the difference between coordinates; if difference b/t coordinates is negative, gene coded on minus strand
GeneProductsDF$strand<-lapply(GeneProductsDF$CoordDiff, StrandType) #Identifying if the gene is coded on plus or minus strand (based on direction of coordinates/coordinates differences)
GeneProductsDF$CoordDiff<-NULL
GeneProductsDF$V1<-NULL

CorrectedCoordinates<-NULL
for (i in 1:nrow(GeneProductsDF)) {
  myrow<-GeneProductsDF[i,]
  if (myrow$strand == "+") {myrow$start_new<-myrow$start  #for genes coded on the plus strand; keep coordinates the same (start= start; end=end)
  myrow$end_new<-myrow$end}
  else {myrow$start_new<-myrow$end #for genes coded on the minus strand, switch start and end coordinates so that they are in the orientation required by the guide finder program
  myrow$end_new<-myrow$start}
  CorrectedCoordinates<-rbind(CorrectedCoordinates,myrow)
}

CorrectedCoordinates$start<-NULL
CorrectedCoordinates$end<-NULL
names(CorrectedCoordinates)<-c("UniqueID","strand", "start", "end")

```

__Writing output file__ 
Converting to matrix in order to write .csv file.  
This file contains the new, modified coordinates (adjusted into proper format that the guide finder program requires)
```{r}
CorrectedCoordinates<-as.matrix(CorrectedCoordinates)
write.csv(CorrectedCoordinates, file = "DraftGenome_NewCoordinates.csv", row.names = F )
```

